\huge Gap formation and stability in non-isothermal disks\\ \Large [2014 CITA Summer Student Program]

### 2014-07-23 poster/poster.tex
poster making, 70%

\title{\huge Gap formation and stability in non-isothermal
  disks\\ \Large [2014 CITA Summer Student Program]}
\author{Min-Kai Lin and Robert Les}
\institute[CITA]{Canadian Institute for Theoretical Astrophysics, 60
  St George Street, Toronto, M5S 3H8, Canada}
%% \date[Sep. 8th, 2009]{Sep. 8th, 2009}

@@

    % ---------------------------------------------------------%
    % Set up a column 
    \begin{column}{.49\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth}  % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content            
            \begin{block}{{\Large Introduction}}
              \justifying
                  {\large              
                    {\bf
                    }
                  }
            \end{block}
            \vfill

            \begin{block}{{\Large Disk-planet model with simple
                  cooling}}
              \justifying
             We numerically evolve the 2D fluid equations including a
             gap-opening giant planet, treated as an external
             potential on fixed circular orbit. We include in the
             energy equation  
        a simple source term in the form 
	\begin{align*}
	\frac{\partial e}{\partial t} = \cdots + \frac{e-e_{t=0}}{t_\mathrm{cool}},
	\end{align*}
        and $t_\mathrm{cool} = \beta \Omega_k^{-1}$, where $\Omega_k$ is the Keplerian frequency. We choose $\beta$ indirectly 
        through $\tilde{\beta}$ such that
        \begin{align*}
        t_\mathrm{cool, gap\,edge} = \tilde{\beta} t_\mathrm{lib, gap\,edge},
        \end{align*}
        where $t_\mathrm{lib, gap\,edge}$ is the time interval between successive encounters of a fluid element, 
        located at the gap edge, and the planet's azimuth. Thus $\tilde{\beta}\ll 1$ implies that such a fluid returns to its initial temperature
        before passing by the planet again. 
        \end{block}
        \vfill

        \begin{block}{{\Large Planet gaps as a function of cooling
              rate}}
          \justifying
          % {\bf generalized vortensity profile due to planet for
          %   different beta. also density profiles.}

          \begin{figure}
            \centering
%            \hfill
%            \begin{minipage}{0.49\textwidth}
              \includegraphics[width=.49\textwidth]{Posterfig_Density}\\
%           \end{minipage}
%            \hfill
%            \begin{minipage}{0.49\textwidth}
              \includegraphics[width=.49\textwidth]{Posterfig_Vortensity}
%            \end{minipage}
%            \hfill
          \end{figure}

          %\begin{figure}
          %  \centering
          %  \includegraphics[width=0.5\textwidth]{Posterfig_HR}
          %\end{figure}

@@

            \begin{block}{{\Large Linear stability}}
              \justifying
                  {\bf plot of growth rate v.s m for different beta,
                    or table. 2D density plot showing development of
                    vortex, for different beta}

                   \begin{figure}
                    \centering
                    \hfill
                    \begin{minipage}{0.45\textwidth}
                      \includegraphics[width=\textwidth]{Posterfig_lowb}
                    \end{minipage}
                    \hfill
                    \begin{minipage}{0.45\textwidth}
                      \includegraphics[width=\textwidth]{Posterfig_medb}
                    \end{minipage}
                    \hfill
                  \end{figure}

                   \begin{table}
                     \hfill
                     \begin{minipage}{\textwidth}
                       \begin{tabularx}{\textwidth}{l*{10}{R}} %\toprule \addlinespace[10pt]
                         \multicolumn{11}{c}{$\tilde{\beta}=0.1$} \\ \midrule
                         m                    & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10  \\ 
                         $\gamma10^2/\Omega(r_o)$ & 6.56 & 6.82 & 6.73 & 5.78 & 6.00 & 6.38 & 5.97 & 5.62 & 4.61 & 3.36   \\ \bottomrule
                       \end{tabularx}
                     \end{minipage}

                     \vspace{1in}

                     \begin{minipage}{\textwidth}
                       \begin{tabularx}{\textwidth}{l*{5}{R}} %\toprule \addlinespace[10pt]
                         \multicolumn{6}{c}{$\tilde{\beta}=1.0$} \\ \midrule
                         m                    & 1 & 2 & 3 & 4 & 5  \\ 
                         $\gamma10^2/\Omega(r_o)$ & 1.27 & 1.28 & 1.35 & 1.01 & 0.61   \\ \bottomrule
                       \end{tabularx}
                     \end{minipage}
                     \hfill
                   \end{table}

@@

            \begin{block}{\Large{Linear stability Cont.}}
              \justifying
              \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{Posterfig_Growth}
              \end{figure}
            \end{block}

@@

            \begin{block}{\Large{Non-linear vortex evolution}}
              \justifying
                  {\bf `lifetime' plot of m=1 for diff beta, snapshots
                    (before, during, after vortex decay). quote rossby
                  number when vortices in quasi-steady state}

                  \begin{figure}
                    \centering
                    \includegraphics[width=\textwidth]{Posterfig_Lifetime_b}
                  \end{figure}

            \end{block}
            \vfill

            \begin{block}{{\Large Conclusions so far}}\justifying
              {\large{\bf

>>>>

\title{\huge Gap formation and stability in non-isothermal
  disks\\ \Large [2014 CITA Summer Student Program]}
\author{Min-Kai Lin and Robert Les$^\dagger$}
\institute[CITA]{Canadian Institute for Theoretical Astrophysics, 60
  St George Street, Toronto, M5S 3H8, Canada\\{\footnotesize$^\dagger$student: all
  simulations, data analysis and figures}}

%% \date[Sep. 8th, 2009]{Sep. 8th, 2009}

@@

    % ---------------------------------------------------------%
    % Set up a column 
    \begin{column}{.49\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth}  % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content            
            \begin{block}{{\Large Introduction}}
              \justifying
                  {\large              
                    {\bf
                      Unstable disk-planet, leading to the
                      formation of large-scale vortices, has been
                      suggested to explain observations of transition
                      disks with lopsided dust
                      distributions. While theoretical disk-planet
                      models employed locally isothermal disks, 
                      the original linear theory of this vortex
                      instability was developed for adiabataic disks. We
                      generalize the study of planetary gap edge
                      instabilities to non-isothermal disks by
                      including an energy equation with variable
                      cooling rate, and study its effect on the
                      formation and evolution of gap-edge vortices. 
                    }
                  }
            \end{block}
            \vfill

            \begin{block}{{\Large Disk-planet model with simple
                  cooling}}
              \justifying
             We numerically evolve the 2D fluid equations including a
             gap-opening giant planet, treated as an external
             potential on fixed circular orbit. We include in the
             energy equation  
        a simple source term in the form 
	\begin{align*}
	\frac{\partial e}{\partial t} = \cdots + \frac{e-e_{t=0}}{t_\mathrm{cool}},
	\end{align*}
        and $t_\mathrm{cool} = \beta \Omega_k^{-1}$, where $\Omega_k$
        is the Keplerian frequency. The energy density
        $e=p/(\gamma-1)$ where $p$ is the pressure field and
        $\gamma=1.4$ is the adiabatic index. We choose $\beta$ indirectly 
        through $\tilde{\beta}$ such that
        \begin{align*}
        t_\mathrm{cool, gap\,edge} = \tilde{\beta} t_\mathrm{lib, gap\,edge},
        \end{align*}
        where $t_\mathrm{lib, gap\,edge}$ is the time interval between
        successive encounters of a fluid element 
        located at the gap edge, and the planet's azimuth. Thus $\tilde{\beta}\ll 1$ implies that such a fluid returns to its initial temperature
        before passing by the planet again. 
        \end{block}
        \vfill

        \begin{block}{{\Large Planetary gap profiles and cooling
              rate}}
          \justifying
          %Profiles during initial stages of gap formation $\to$
          \begin{minipage}[t]{0.49\textwidth}
            \begin{figure}
              %            \hfill
              \includegraphics[width=.99\textwidth,clip=true,trim=0.5cm
                2cm 0cm 1cm]{Posterfig_Density}\\
              \includegraphics[width=.99\textwidth,clip=true,trim=0.5cm
                0.5cm 0cm 1cm]{Posterfig_Vortensity}
            \end{figure}
          \end{minipage}
          \begin{minipage}[t]{0.49\textwidth}
            \begin{itemize}
            \item Top panel: shallower gaps with increasing $t_\mathrm{cool}$ 
              but no strong dependence 
            \item Lower panel shows the generalized vortensity
              \begin{align*}
                \tilde{\eta} = \frac{\omega_z}{\Sigma}\times S^{-2/\gamma},
              \end{align*}
              where $\omega_z$ is the vertical vorticity and $S\equiv
              p/\Sigma^\gamma$ is the entropy. 
            \item Extrema in $\tilde{\eta}$ associated with
              instability (Lovelace et al. 1999, Li et al. 2000)
            \item Entropy-production across planet-induced shocks
              weakens extrema in $\tilde{\eta}$, so
             increasing $t_\mathrm{cool}$ may stabilize the
              system by making gap edges less sharp
%              \item A sufficiently deep minima in $\tilde{\eta}$ can
%                cause instability
%              \item 
            \end{itemize}
          \end{minipage}
          %\begin{figure}
          %  \centering
          %  \includegraphics[width=0.5\textwidth]{Posterfig_HR}
          %\end{figure}

@@

        \begin{block}{{\Large Numerical linear stability simulations}}
          \justifying
          We switch off the planet potential after it is fully
          introduced and a partial gap is already formed. An azimuthal
          average is performed, and the system is subject to random
          density perturbations.\\   
          \vspace{-1.5cm}
%                  {\bf plot of growth rate v.s m for different beta,
%                    or table. 2D density plot showing development of
%                    vortex, for different beta}                   
                    \hfill
                    \begin{minipage}[t]{0.49\textwidth}
                      \begin{figure}
                        \centering
                        \includegraphics[width=.95\textwidth,clip=false,trim=0.01cm
                          0cm .1cm 0cm]{Posterfig_lowb}
                      \end{figure}
                    \end{minipage}
                                        %\hfill
                    \begin{minipage}[t]{0.49\textwidth}
                      \begin{figure}
                        \includegraphics[width=.95\textwidth,clip=false,trim=0cm
                        0cm .1cm 0cm]{Posterfig_medb}
                      \end{figure}
                    \end{minipage}    
                    \begin{itemize}
                    \item Slowering cooling rate increases
                      \emph{linear} stability and favors lower-$m$ modes
                    \end{itemize}
%                   \begin{table}
%                     \hfill
%                     \begin{minipage}{\textwidth}
%                       \begin{tabularx}{\textwidth}{l*{10}{R}} %\toprule \addlinespace[10pt]
%                         \multicolumn{11}{c}{$\tilde{\beta}=0.1$} \\ \midrule
%                         m                    & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10  \\ 
%                         $\gamma10^2/\Omega(r_o)$ & 6.56 & 6.82 & 6.73 & 5.78 & 6.00 & 6.38 & 5.97 & 5.62 & 4.61 & 3.36   \\ \bottomrule
%                       \end{tabularx}
%                     \end{minipage}
%
%                     \vspace{1in}
%
%                     \begin{minipage}{\textwidth}
%                       \begin{tabularx}{\textwidth}{l*{5}{R}} %\toprule \addlinespace[10pt]
%                         \multicolumn{6}{c}{$\tilde{\beta}=1.0$} \\ \midrule
%                         m                    & 1 & 2 & 3 & 4 & 5  \\ 
%                         $\gamma10^2/\Omega(r_o)$ & 1.27 & 1.28 & 1.35 & 1.01 & 0.61   \\ \bottomrule
%                       \end{tabularx}
%                     \end{minipage}
%                     \hfill
%                   \end{table}

@@

 %           \begin{block}{\Large{Linear stability Cont.}}
 %             \justifying
 %             \begin{figure}
 %               \centering
 %               \includegraphics[width=\textwidth]{Posterfig_Growth}
 %             \end{figure}
 %           \end{block}
 %
 %           \vfill

            \begin{block}{\Large{Non-linear evolution of gap edge
                  vortices}}
              \justifying
%                  {\bf `lifetime' plot of m=1 for diff beta, snapshots
%                    (before, during, after vortex decay). quote rossby
%                  number when vortices in quasi-steady state}

            \begin{block}{\Large{Vortex life-time and cooling rate}}
              \justifying

                  \begin{figure}
                    \centering
                    \includegraphics[width=\textwidth]{lifetime_regime1}\\
                    \includegraphics[width=\textwidth]{lifetime_regime2}
                  \end{figure}
                  \end{block}

@@

            \begin{block}{{\Large Discussion}}\justifying
              {\large{\bf
